---
title: "Individual Assignment"
output:
  html_document: default
---

```{r, include=FALSE}
library(knitr)
library(readxl)
library(forecast)
library(tseries)
library(vars)
library(stargazer)
library(dplyr)
library(ggplot2)
library(tidyr)
```

Let's load the dataset 'online_retailer' and delete the weeks 133 to 137 as no revenue is recorded for those lasts weeks:

```{r}
data <- read_excel("/Users/coralineduval/Downloads/online_retailer.xlsx",sheet='retail_brand', na='NA')
data=slice(data, 1:132)
head(data)
```


I will focus on the type of adds (mobile, display, paid search) and the different social media conversations around the brand.

```{r}
ts.plot(data$revenues, col="blue", main="Revenues")
ts.plot(data$Mobile_ads, col="darkgreen", main="Mobile ads")
ts.plot(data$Display_ads, col="red", main="Display ads")
ts.plot(data$Paid_GGL, col="maroon", main="Paid search ads") 
```


I will now plot the number of conversation for each of the 4 topics over the weeks : 

```{r}
agg <- data %>%
   filter(! is.na(earned_ads)) %>%
   pivot_longer(cols = earned_went_pruchased:earned_love, names_to = "earned", values_to="Value")
ggplot(agg, aes(x = week_beg, y = Value, group = earned, color = earned)) + geom_line()+ labs(title='Number of conversation on social media by topic', y='Number of conversations',x='Week')
``` 

The current budget allocation between the three type of ads is:

```{r}
#Current budget allocation

cost_mobile<-sum(data$Mobile_ads)
cost_display<-sum(data$Display_ads)
cost_paid<-sum(data$Paid_GGL)
cost_total <- cost_mobile+cost_display+cost_paid

costshare_mobile<-cost_mobile/cost_total
costshare_display<-cost_display/cost_total
costshare_paid<-cost_paid/cost_total

# Ingredients for the pie-chart 
slices_actual<-c(costshare_mobile, costshare_display,costshare_paid )
lbls_actual<-c("Mobile ads", "Display ads", "Paid search ads")
pct_actual<-round(slices_actual*100)
lbls_actual<-paste(lbls_actual, pct_actual)          # add data to labels
lbls_actual<-paste(lbls_actual, "%", sep="")  # add % sign to labels

# Get the pie-chart
pie(slices_actual, labels=lbls_actual, col=rainbow(length(lbls_actual)), main="Actual Budget Allocation" )


```

Then, the log of the variables chosen for this model will be taken to balance out the high volatility observed above.

```{r}
data$Lrevenues<- log(data$revenues+1)
data$LMobile_ads<- log(data$Mobile_ads+1)
data$LDisplay_ads<- log(data$Display_ads+1)
data$LPaid_GGL<- log(data$Paid_GGL+1)
data$Learned_General<- log(data$earned_General+1)
data$Learned_ads<- log(data$earned_ads+1)
data$Learned_love<- log(data$earned_love+1)
data$Learned_went_pruchased<- log(data$earned_went_pruchased+1)
```

## 3. VAR Model Estimation

### 3.1 Pre-estimation Diagnostics

#### 3.1.2 Plot the Data to Inspect Trend

To see if the variables are evolving overtime, the ACF and PACF will be plotted for each variable.   

```{r}
revenues <- ts(data$revenues, frequency = 52, start = c(1, 1))
mobile <- ts(data$Mobile_ads, frequency = 52, start = c(1, 1))
display <- ts(data$Display_ads, frequency = 52, start = c(1, 1))
paid <- ts(data$Paid_GGL, frequency = 52, start = c(1, 1))
eg <- ts(data$earned_General, frequency = 52, start = c(1, 1))
ea <- ts(data$earned_ads, frequency = 52, start = c(1, 1))
el <- ts(data$earned_love, frequency = 52, start = c(1, 1))
ewp <- ts(data$earned_went_pruchased, frequency = 52, start = c(1, 1))
ggtsdisplay(revenues)
ggtsdisplay(mobile)
ggtsdisplay(display)
ggtsdisplay(paid)
ggtsdisplay(eg)
ggtsdisplay(ea)
ggtsdisplay(el)
ggtsdisplay(ewp)

```


#### 3.1.3 Unit Root Testing

The ACF and PACF graphs seen to suggest that most of the variables are stationary. 
I will run a adf test with the null hypothesis "the variable is not stationary". 
Beyond plotting, we can run some hypothesis testing to confirm that. There are multiple tests *adf.test*, *pp.test*, *kpss.test*. For the adf.test and pp.test, if the p-value is smaller than significance level then we can reject the null hypothesis that the variable is not stationary; kpss.test is the reverse, the null hypothesis is that it is stationary. 



```{r}
Lrevenues <- ts(data$Lrevenues, frequency = 52, start = c(1, 1))
Lmobile <- ts(data$LMobile_ads, frequency = 52, start = c(1, 1))
Ldisplay <- ts(data$LDisplay_ads, frequency = 52, start = c(1, 1))
Lpaid <- ts(data$LPaid_GGL, frequency = 52, start = c(1, 1))
Leg <- ts(filter(data,! is.na(Learned_General))$Learned_General, frequency = 52, start = c(2, 23))
Lea <- ts(filter(data,! is.na(Learned_ads))$Learned_ads, frequency = 52, start = c(2, 23))
Lel <- ts(filter(data,! is.na(Learned_love))$Learned_love, frequency = 52, start = c(2, 23))
Lewp <- ts(filter(data,! is.na(Learned_went_pruchased))$Learned_went_pruchased, frequency = 52, start = c(2, 23))
adf.test(Lrevenues)
adf.test(Lmobile)
adf.test(Ldisplay)
adf.test(Lpaid)
adf.test(Leg)
adf.test(Lea)
adf.test(Lel)
adf.test(Lewp)


```

Only Lewp seem to have p>0.05, so we can accept the null hypothesis that the variable is not stationnary and take the first difference to make it stationary. I then run again a adf test to make sure it is now stationnary: 


```{r}
DLewp <-diff(Lewp, differences = 1)
adf.test(DLewp) #first-differenced series
```


### 3.2 Construct two VAR model

I will construct the VAR model for the ads and for the social media conversations topics.

For the online ads (starting at (1,1) as there is not differentiated variables):

```{r warning=FALSE}
#Build a dataset for VAR model
data.ts.d1 <- window(cbind( Lmobile, Ldisplay,Lpaid,Lrevenues), start = c(1, 1))

varp <- VAR(data.ts.d1, ic="AIC", lag.max=1, type="const")

lmp <- varp$varresult

stargazer(lmp$Lrevenues, lmp$Lmobile, lmp$Ldisplay,lmp$Lpaid, column.labels = c('revenues', 'mobile', 'display', 'paid'), type = "text", dep.var.labels.include = FALSE )
```
For the social media conversations topics (starting at (2,24) because of the data missing for the first 75 weeks and because of the differentiation of DLewp):

```{r warning=FALSE}
#Build a dataset for VAR model
data.ts.d2 <- window(cbind( Leg ,Lea ,Lel,DLewp,Lrevenues), start = c(2, 24))

varp2 <- VAR(data.ts.d2, ic="AIC", lag.max=1, type="const")

lmp2 <- varp2$varresult

stargazer(lmp2$Lrevenues,lmp2$Leg ,lmp2$Lea ,lmp2$Lel,lmp2$DLewp, column.labels = c('revenues', 'general', 'ads', 'love','went purchased'), type = "text", dep.var.labels.include = FALSE )
```


#### Check the residuals

I am checking for both models the residuals' normality and the autocorrelation : 


```{r}

revenues.residuals <- data.frame(residuals(varp))$Lrevenues
revenues.residuals <- ts(revenues.residuals, frequency = 52, start = c(1, 1))
round(mean(revenues.residuals),4)
autoplot(revenues.residuals)

revenues2.residuals <- data.frame(residuals(varp2))$Lrevenues
revenues2.residuals <- ts(revenues2.residuals, frequency = 52, start = c(1, 1))
round(mean(revenues2.residuals),4)
autoplot(revenues2.residuals)

```

Both seem quite reasonable as the mean is around 0.

I can now look at the short vs long-term effects of the ads spendings and the topics in social media conversation and the optimal allocation between the three possible online ads. 


### 3.3 IRF Analysis

Plotting the IRF plots for the ads : 


```{r}
irfs <- irf(varp, impulse = c('Lmobile', 'Ldisplay','Lpaid'), response = 'Lrevenues', 
            runs = 100, n.ahead = 7 , ortho = TRUE, ci=0.95)

plot(irfs)
```
Plotting the IRF plots for the topics in social media conversations : 

```{r}
irfs2 <- irf(varp2, impulse = c('Leg' ,'Lea' ,'Lel','DLewp'), response = 'Lrevenues', 
            runs = 100, n.ahead = 7 , ortho = TRUE, ci=0.95)

plot(irfs2)
```


#### Immediate and Long-Term Effects

In the following part I compute the t-statistics for the IRF coefficient to see if it is significant (>1) or not.


```{r}
#Make a table to summarize IRF coefficients and their confidence intervals

irf.table.ci <- round(data.frame(period = seq(1, 8), 
                           response.Mobile = irfs$irf$Lmobile, 
                           Mobile.lower = irfs$Lower$Lmobile, 
                           Mobile.upper = irfs$Upper$Lmobile, 
                           response.Display = irfs$irf$Ldisplay, 
                           display.lower = irfs$Lower$Ldisplay, 
                           display.upper = irfs$Upper$Ldisplay,
                           response.Paid = irfs$irf$Lpaid, 
                           paid.lower = irfs$Lower$Lpaid, 
                           paid.upper = irfs$Upper$Lpaid),4)
colnames(irf.table.ci) <- c('Period', 'Lmobile', 'Lmobile Lower', 'Lmobile Upper','Display', 
                            'Ldisplay Lower', 'Ldisplay Upper','Paid', 
                            'Lpaid Lower', 'Lpaid Upper')

knitr::kable(irf.table.ci)
```

I then apply the t>1 criteria and calculate the long term elasticities for each type of ad spending:

```{r}
#Mobile
result_irf_mobile<-matrix(nrow = 8, ncol = 1)

for (i in 1:8) {
  se <- (irfs$Upper$Lmobile[i]-irfs$Lower$Lmobile[i])/(2*1.96)
  t_irf_mobile<- irfs$irf$Lmobile[i]/se
   
   if (t_irf_mobile>1) {
    result_irf_mobile[i] <- irfs$irf$Lmobile[i]
   } else {
      result_irf_mobile[i] <-0
      }
}
result_irf_mobile
lr_mobile <- sum(result_irf_mobile)
lr_mobile
```

An increase in mobile ads spendings will have the largest positive impact on revenues in the first and second period.
An 1% increase in mobile advertising spending will increase the retailer revenues by 0.23% in the long run.

```{r}
#Display spending

result_irf_display<-matrix(nrow = 8, ncol = 1)

for (i in 1:8) {
  se <- (irfs$Upper$Ldisplay[i]-irfs$Lower$Ldisplay[i])/(2*1.96)
  t_irf_display<- irfs$irf$Ldisplay[i]/se
   
   if (t_irf_display>1) {
    result_irf_display[i] <- irfs$irf$Ldisplay[i]
   } else {
      result_irf_display[i] <-0
      }
}
result_irf_display
lr_display <- sum(result_irf_display)
lr_display
```

An increase in display ads spendings will have a positive impact (while small) on revenues in the first period.
An 1% increase in display advertising spending will increase the retailer revenues by 0.02% in the long run.

```{r}
#Paid

result_irf_paid<-matrix(nrow = 8, ncol = 1)

for (i in 1:8) {
  se <- (irfs$Upper$Lpaid[i]-irfs$Lower$Lpaid[i])/(2*1.96)
  t_irf_paid<- irfs$irf$Lpaid[i]/se
   
   if (t_irf_paid>1) {
    result_irf_paid[i] <- irfs$irf$Lpaid[i]
   } else {
      result_irf_paid[i] <-0
      }
}
result_irf_paid
lr_paid <- sum(result_irf_paid)
lr_paid
```

An increase in paid seraches ads spendings will have a large positive impact on revenues in the first and second period.
An 1% increase in paid searches advertising spending will increase the retailer revenues by 0.20% in the long run.

### 3.4 Optimal allocation of spending

I will now compute the optimal allocation between the three types of ads.

```{r}

#Get the coefficients from IRF results
beta_mobile<-lr_mobile
beta_display<-lr_display
beta_paid<-lr_paid

#The sum of all elasticities 
beta_all<-beta_mobile+beta_display+beta_paid

#Optimal resource allocation
optim_mobile<-beta_mobile/beta_all
optim_display<-beta_display/beta_all
optim_paid<-beta_paid/beta_all
```

Here is the resulting budget allocation :
```{r}
## Pie-chart ingredients 
optimal_spend<-c(optim_mobile,optim_display, optim_paid)
optimal_spend=round(optimal_spend, digits=5)
optimal_spend

slices_optim<-c(optim_mobile,optim_display, optim_paid)
lbls_optim<-c("Mobile ads", "Display ads", "Paid search ads")
pct_optim<-round(slices_optim*100)
lbls_optim<-paste(lbls_optim, pct_optim)   # paste variable names to data labels 
lbls_optim<-paste(lbls_optim, "%", sep="") # add % sign to labels

# Get the pie-chart
pie(slices_optim, labels=lbls_optim, col=rainbow(length(lbls_optim)), main="Optimal Budget Allocation" )

```



